### 2015-10-18 abstract.tex
some proofreading fixes and improvements

\begin{abstract}
The \emph{period enforcer} algorithm for self-suspending real-time tasks is a technique for suppressing the ``back-to-back'' scheduling penalty associated with deferred execution. Originally proposed in 1991, the algorithm has attracted renewed interest in recent years. This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~with current techniques, schedulability analysis of the period enforcer algorithm requires a proper transformation, that is subject to exponential time complexity, of the task set to satisfy the assumptions; and \textbf{(iii)} the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
\end{abstract}

>>>>

\begin{abstract}
The \emph{period enforcer} algorithm for self-suspending real-time tasks is a technique for suppressing the ``back-to-back'' scheduling penalty associated with deferred execution. Originally proposed in 1991, the algorithm has attracted renewed interest in recent years. This note revisits the algorithm in the light of recent developments in the analysis of self-suspending tasks, carefully re-examines and explains its underlying assumptions and limitations, and points out three observations that have not been made in the literature to date: \textbf{(i)}~period enforcement is not strictly superior (compared to the base case without enforcement) as it can cause deadline misses in self-suspending task sets that are schedulable without enforcement; \textbf{(ii)}~to match the assumptions underlying the analysis of the period enforcer, a schedulability analysis of self-suspending tasks subject to period enforcement requires a task set  transformation that, with current techniques, is subject to exponential time complexity; and \textbf{(iii)}~the period enforcer algorithm is incompatible with all existing analyses of suspension-based locking protocols, and can in fact cause ever-increasing suspension times until a deadline is missed.
\end{abstract}

### 2015-10-18 locking.tex
some proofreading fixes and improvements

Since $\tau_2$'s first job requests the shared resource only after two time units of execution, it is blocked by $\tau_1$'s critical section, which commenced at time $1$. At time $3$, $\tau_1$ releases the shared resource and $\tau_2$ consequently resumes (i.e., $a^2_{2,1} = 3$). According to the period enforcer rules~\cite{Raj:suspension1991}, the second segment is immediately eligible because, according to Equation~\ref{eq:ET-def},
\begin{align*}
	ET_{2,1}^2 & = \max\left(ET_{2,0}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,1})\right) =\max(-T_2 + T_2,\ 3) = 3.
\end{align*}
(Recall that $ET_{2,0}^2 = -T_2$, and interpret $\mathit{busy}(\tau_2, a^2_{2,1})$ with respect to $\tau_2$'s processor.)

At time $7$, the second job of $\tau_2$ is released. Its first segment ends at time $9$. However, its second segment is not eligible to be scheduled before time $10$ since $ET_{2,2}^2 \geq ET_{2,1}^2 + T_2 = 3 + 7 = 10$. At time $9$, the second job of $\tau_1$, released at time $8$, can thus lock the shared resource without contention. Consequently, when $\tau_2$'s request for the shared resource takes effect at time $10$, the resource is no longer available and $\tau_2$ must wait until time $a^2_{2,2} = 11$ before it can proceed in its execution. We thus have
\begin{align*}
	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,2})\right) =\max(10,\ 11) = 11.
\end{align*}

>>>>

Since $\tau_2$'s first job requests the shared resource only after two time units of execution, it is blocked by $\tau_1$'s critical section, which commenced at time $1$. At time $3$, $\tau_1$ releases the shared resource and $\tau_2$ consequently resumes (i.e., $a^2_{2,1} = 3$). According to the period enforcer rules~\cite{Raj:suspension1991}, the second segment is immediately eligible because, according to Equation~\ref{eq:ET-def} (in Section~\ref{sec:unschedulable}),
\begin{align*}
	ET_{2,1}^2 & = \max\left(ET_{2,0}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,1})\right) =\max(-T_2 + T_2,\ 3) = 3.
\end{align*}
(Recall that $ET_{2,0}^2 = -T_2$, and interpret $\mathit{busy}(\tau_2, a^2_{2,1})$ with respect to $\tau_2$'s processor.)

At time $7$, the second job of $\tau_2$ is released. Its first segment ends at time $9$. However, its second segment is not eligible to be scheduled before time $10$ since $ET_{2,2}^2 \geq ET_{2,1}^2 + T_2 = 3 + 7 = 10$. At time $9$, the second job of $\tau_1$, released at time $8$, can thus lock the shared resource without contention. Consequently, when $\tau_2$'s request for the shared resource takes effect at time $10$, the resource is no longer available and $\tau_2$ must wait until time $a^2_{2,2} = 11$ before it can proceed to execute. We thus have
\begin{align*}
	ET_{2,2}^2 & = \max\left(ET_{2,1}^2 + T_2,\ \mathit{busy}(\tau_2, a^2_{2,2})\right) =\max(10,\ 11) = 11.
\end{align*}

### 2015-10-18 untitled.tex
some proofreading fixes and improvements

In contrast to the first job, the second job of task $\tau_2$ (which is released at time $11$) is affected by period enforcement. The first segment of the second job arrives at time $a^1_{2,2} = 11$, is interfered with for one time unit, and suspends at time $13$. The segment second of the second job hence resumes only at time $a^2_{2,2} = 19$. Thus, we have
\begin{align*}
	ET_{2,2}^1 & = \max\left(0 + 11,\ \mathit{busy}(\tau_2, 11)\right) = 11  \text{ and }
\\
	ET_{2,2}^2 & = \max\left(9 + 11,\ \mathit{busy}(\tau_2, 19\right) ) = 20.
\end{align*}
According to the rules of the period enforcer algorithm, the processor therefore remains idle at time $19$ because the segment is not eligible to execute until time $ET_{2,2}^2 = 20$. However, at time $20$, the third job of $\tau_1$ is released. As a result, the second job of $\tau_2$ incurs additional interference and misses its deadline at time $22$.

@@

One may consider to enrich the period enforcer with the following power: when the processor becomes idle, a task immediately becomes eligible to execute regardless of its eligibility time. However, even with this exension, the above example remains valid by introducing one additional lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from time $3$ to time $9$ and time $13$ to time $20$) and $T_3=D_3=100$. With task $\tau_3$, the processor is always busy from time $0$ to time $23$ and consequently $\tau_2$ still misses its deadline at time $22$.

>>>>

In contrast to the first job, the second job of task $\tau_2$ (which is released at time $11$) is affected by period enforcement. The first segment of the second job arrives at time $a^1_{2,2} = 11$, is interfered with for one time unit, and suspends at time $13$. The  second segment of the second job hence resumes only at time $a^2_{2,2} = 19$. Thus, we have
\begin{align*}
	ET_{2,2}^1 & = \max\left(0 + 11,\ \mathit{busy}(\tau_2, 11)\right) = 11  \text{ and }
\\
	ET_{2,2}^2 & = \max\left(9 + 11,\ \mathit{busy}(\tau_2, 19\right) ) = 20.
\end{align*}
According to the rules of the period enforcer algorithm, the processor therefore remains idle at time $19$ because the segment is not eligible to execute until time $ET_{2,2}^2 = 20$. However, at time $20$, the third job of $\tau_1$ is released. As a result, the second job of $\tau_2$ incurs additional interference and misses its deadline at time $22$.

@@

One may consider to enrich the period enforcer with the following power: when the processor becomes idle, a task immediately becomes eligible to execute regardless of its eligibility time. However, even with this extension, the above example remains valid by introducing one additional lowest priority task $\tau_3$ with execution time $C_3=13$ (to be executed from time $3$ to time $9$ and time $13$ to time $20$) and $T_3=D_3=100$. With task $\tau_3$, the processor is always busy from time $0$ to time $23$ and consequently $\tau_2$ still misses its deadline at time $22$.
